

\section{Flow Engineering Formulation}

\subsection{Notation}

\begin{table*}[!t]
\centering
\small
\begin{tabular}{c|l}
\hline
$S$ & Set of all switches \\
$S_{ToR}$ & Set of all ToR switches \\
$\tau_{u}, \forall u \in S$ & Number of flow entries to install in the switch $u$ \\
$E$ & Set of all physical links (between two adjacent switches) \\
$C_e, \forall e \in E$ & capacity of individual links \\
$F_{uv}$ & set of all flows from $u$ to $v$ where $u,v \in S_{ToR}$ \\
$P_{uv}$ & set of paths from switch $u$ to switch $v$ \\
$K _{uv}$ & number of paths from switch $u$ to $v$ where $u,v \in S_{ToR}$ \\
$P^k _{uv}$ & set of links of $k^{th}$ path from device $u$ to $v$ where $u,v \in S_{ToR}$ \\
$T ^f _{uv}$ & traffic volume from $u$ to $v$ of flow $f$ where $u,v \in S_{ToR}$ \\
$util$ & maximum link utilization \\
$I ^{fk} _{uv}$ & indicator variable denoting that flow $f$ from $u$ to $v$  takes the $k^{th}$ path \\
$D_{u} \forall u \in S$ & Number of low priority rules in the switch $u$ \\
\hline
%\tabincell{c}{HP Procurve \\8212zl} & 666Mhz & 256MB & $\approx$1000 & Modularity \\ \hline
%\hline
\end{tabular}
\topcompactcaption{Table of Notations}{\label{notations}}
\end{table*}

\iffalse
\begin{itemize}
\item $S$ - Set of all switches
\item $S_{ToR}$ - Set of all ToR switches
\item $\tau_{u}, \forall u \in S$ - Maximum number of flow entries in the switch $u$
\item $E$ - Set of all physical links (between two adjacent devices)
\item $C_e, \forall e \in E$ - capacity of individual links
\item $F_{uv}$ - set of all flows from $u$ to $v$ where $u,v \in S_{ToR}$
\item $P_{uv}$ - Set of paths from device $u$ to device $v$
\item $K _{uv}$ - Number of paths from device $u$ to $v$ where $u,v \in S_{ToR}$ 
\item $P^k _{uv}$ - Set of links of $k^{th}$ path from device $u$ to $v$ where $u,v \in S_{ToR}$
\item $T ^f _{uv}$ - Traffic volume from $u$ to $v$ of flow $f$ where $u,v \in S_{ToR}$
\item $util$ - maximum link utilization
\item $I ^{fk} _{uv}$ - Indicator variable denoting that flow $f$ from $u$ to $v$  takes the $k^{th}$ path.
%\item $entries$ - maximum number of flow rules in any switch
\end{itemize}
\fi

\subsection{Step 1 - Minimize ``utilization''}
	In first step we minimize the maximum link utilization. We represent a network as a graph \textit{G = (V,E)}, where each vertex is either a server rack or switch and each edge is a physical link. Let $T ^f _{uv}$ be the traffic volume between $(u,v)$ of flow \textit{f}, $P^k _{uv}$ be the set of links on the $k^{th}$ path between (\textit{u,v}). Following are our two constraints. First, the total traffic volume from an edge should not exceed its $C_e$ times the link utilization limit, $util$. We use a binary indicator variable  $I ^{fk} _{uv}$ to show whether the flow $f$ is on path $k$ or not.

	
\begin{equation}
    \sum _{u,v \in S_{ToR},~f \in F_{uv},~k  \in P_{uv}~s.t.~e \in P^k _{uv}} T ^f _{uv} *  I^{fk}_{uv} \leq util \times C_e 
\end{equation}
	
	
	
	Second, each flow $f$ should be routed from a single path.

\begin{equation}
    \sum _{k  \in P_{uv}} I^{fk}_{uv} = 1 ,\quad \quad \forall u,v \in S_{ToR},~ f \in F_{uv}
\end{equation}
  

The objective is to minimize the \textit{util}, the maximum link utilization, under the above two constraints. 

\subsection{Step 2 - Minimize ``rule setup latency''}
Our measurement shows that flow installation latency has a linear relation with the burst size of the rules (number of rules to be installed). Flow installation latency can be define as
  \begin{equation*}
            latency_{u} = (a_{u}  + c_{u}* D_{u}) *\tau_{u}
  \end{equation*}
, where $a_{u}$ and $c_{u}$ are constants for switch $u$.

In second step, we minimize the latency of installing rules. Following are three constraints.
 
First,  traffic traversing from an edge $e$ does not exceed the $C_e$ multiplied by 1 $+$ $\delta$ times the $util^*$. $util^*$ is the maximum link utilization we got from the step 1, and  $\delta$ defines an upper limit on maximum link utilization.

\begin{equation}
    \sum _{u,v \in S_{ToR},~f \in F_{uv},~k  \in P_{uv}~s.t.~e \in P^k _{uv}} T ^f _{uv} *  I^{fk}_{uv} \leq (1 + \delta) \times util^* \times C_e 
\end{equation}

  Second, constraint is same as step 1.
  
\begin{equation}
      \sum _{k  \in P_{uv}} I^{fk}_{uv} = 1 ,\quad \quad \forall u,v \in S_{ToR},~ f \in F_{uv}
\end{equation}

  Third, flows passing from a switch $s$ are less than the number of maximum rules in any switch in the network.
\begin{equation}
                       \sum_{u,v \in S_{ToR}, ~f \in F_{uv}, ~k \in P_{uv} ~s.t.~ s\in P^k_{uv} } I^{fk}_{uv} \leq \tau_{u} \quad \quad \forall s \in S
\end{equation}
The objective is to minimize the $latency$ on any switch, under above three constraints.   
  

